From Reversible Programs to \\ Univalent Universes and Back

### 2017-06-30 upi.tex
start working on section 3.  Introduce some macros so that we are not so display driven.  Edit well beyond the wavefront, but 3.1 is mysterious to me, so work on 3.2.

@@

%%%%%
\subsection{$\Pi_{\mathbb{2}}$}

@@

@@

Instead of guessing, there is a special class of univalent fibrations:
\[\begin{array}{rcl}
\{F\} &:& \Sigma_{A:\mathcal{U}} \| F \equiv A \| \rightarrow \mathcal{U} \\
\{F\}(A,|p|) &=& A
\end{array}\]

This construction is always a univalent fibration~\cite{XXX}. We can understand
the basic idea as follows. Take two points $(A,|p|)$ and $(B,|q|)$ in the base
space. We want to compare the space $A \simeq B$ with the space
$(A,|p|) \equiv (B,|q|)$.  In one direction, we have $p : F \equiv A$ and
$q : F \equiv B$ and hence $A \equiv B$ and by univalence $A \simeq B$. In the
other direction \ldots

@@

>>>>

% TODO: give this a better name!
\newcommand{\bracket}[1]{\ensuremath{\{#1\}}}
\newcommand{\ptrunc}[1]{\ensuremath{\| #1 \|}}

\newcommand{\PiTwo}{\ensuremath{\Pi_{\mathbb{2}}}}

@@

%%%%%
\subsection{\PiTwo}{\label{sec:pi2}}

@@

It turns out that there is a special class of fibrations.  For its
definition, we need to first introduce the idea of
\emph{propositional truncation}, and a special map from
paths to equivalences.

\begin{defn}
Given a type $A : \mathcal{U}$, the type
$\ptrunc{A}$ is the propositional truncation of $A$ defined as: 
\begin{itemize}
\item for $x : A$, we have $|x| : \ptrunc{A}$,
\item for any $x,y : \ptrunc{A}$, we have $\gluep : x \equiv y$.
\end{itemize}
\end{defn}
\noindent We can understand $\ptrunc{A}$ as a space which has
$|a|$ for all $a : A$ as its points, but where all such points are
(propositionally) identified via $\gluep$.  In other others,
by continuity, $\ptrunc{A}$ behaves the same as $\mathbb{1}$,
in that all inhabitants of $\ptrunc{A}$ give rise to the same
behaviour in \emph{all} situations.

\begin{defn}[Lemma 2.10.1 of~\cite{HoTT}]
Given two types $A$ and $B$, there is a function
$\mathit{idtoeqv} : (A \equiv B) \to (A \simeq B)$.
\end{defn}

For any type $F$, let us define $\bracket{F}$ via
\[\begin{array}{rcllcl}
\bracket{F} &:& \Sigma_{A:\mathcal{U}} & \ptrunc{F \equiv A} & \rightarrow & \mathcal{U} \\
\bracket{F} & & (A, & |p|) &=& A
\end{array}\]

@@

\jacques{The following is starting to be quite fuzzy}
This construction always gives a univalent fibration~\cite{Christiansen},
whenever the universe $\mathcal{U}$ is univalent. We can understand
the basic idea as follows. Take two points $(A,|p|)$ and $(B,|q|)$ in the base
space. We want to compare the space $A \simeq B$ with the space
$(A,|p|) \equiv (B,|q|)$.  
\begin{itemize}
\item Given $(A,|p|)$ and $(B,|q|)$, we have $p : F \equiv A$ and $q : F \equiv B$.
By symmetry and transitivity of $\equiv$, we have $A \equiv B$; by
$\mathit{ittoequiv}$, $A \simeq B$.
\item Given $A \simeq B$, by univalence of $\mathcal{U}$, we get that
$A \equiv B$.  
\end{itemize}

@@

\subsection{Looking at $\bracket{\bt}$}

We claim that there is a link between $\PiTwo$ (defined in Sec.~\ref{sec:pi2})
and $\bracket{\bt}$.  To see this, we begin by unpacking the points, 1-paths,
and 2-paths of $\bracket{\bt}$.

\paragraph*{Points in $\bracket{\bt}$.} The only term of type $\bracket{\bt}$ is
$(\bt,\|\reflp_{\bt}\|)$ which we call $`\bt$. Formally we can prove that
$\ptrunc{(X,p) \equiv `\bt}$ is inhabited for all $X$ and $p$.
